---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# carpeDIAMS Tutorial

<!-- badges: start -->
<!-- badges: end -->

**carpeDIAMS** is an R package for converting DIA (Data Independent Acquisition) mass spectrometry data into pseudo-DDA (Data Dependent Acquisition) spectra. This package provides functions to extract MS2 information from DIA experiments and create consolidated spectra based on correlation analysis.

## Installation

You can install the development version of carpeDIAMS like so:

```{r}
pak::pak("https://github.com/JohanLassen/carpeDIAMS")
```



```{r}
library(readr)
library(xcms)
library(tibble)
library(purrr)
library(Spectra)
library(data.table)
library(dplyr)
library(tidyr)
library(SummarizedExperiment)
library(rhdf5)
library(arrow)
library(duckdb)
Rcpp::sourceCpp("/faststorage/project/forensics_TOFscreenings/06_integrated_workflow/modules/carpediams/scripts/gemini_upgma.cpp")
source("/faststorage/project/forensics_TOFscreenings/06_integrated_workflow/modules/carpediams/scripts/spectral_filtering_v4_batch.R")
source("/faststorage/project/forensics_TOFscreenings/06_integrated_workflow/modules/shared/carpeDIAMS/R/catch_ms2_functions.R")

sample <- "/faststorage/project/forensics_TOFscreenings/06_integrated_workflow/results/tmp/peak_picked/TX35284_A_01-346381_1_1-G.1_01_18036.rds"
dataset <- "/faststorage/project/forensics_TOFscreenings/06_integrated_workflow/results_rare/results/feature_definitions.rds"

test <- read_rds(sample) |> xcms::chromPeaks() |> as_tibble()

references <- read_rds(dataset)

references |> as_tibble(rownames = "feature")

mz1 <- test |> pull(mz)
mz2 <- references$mzmed
rt1 <- test |> pull(rt)
rt2 <- references$rtmed
rt <- c(rt1, rt2)
mz <- c(mz1, mz2)

groups <- fast_1D_upgma_grouping_cpp(mz, maxDiff = 0.01)
groups2 <- groups

# Make progress bar in for loop
pb <- txtProgressBar(min = 0, max = length(unique(groups)), style = 3)
for (i in unique(groups)){
  idx <- which(groups == i)
  mz_subset <- mz[idx]
  rt_groups <- fast_1D_upgma_grouping_cpp(rt[idx], maxDiff = 10)
  new_groups <- paste0(i, "_", rt_groups)
  groups2[idx] <- new_groups
  setTxtProgressBar(pb, which(unique(groups) == i))
}
close(pb)
groups2

g1 <- groups2[1:length(mz1)]
g2 <- groups2[(length(mz1)+1):length(groups2)]

# find indices of g2 that exist in g1
indices <- which(g2 %in% g1)
references_subset <- references[indices, ]
references$group <- g2
test$group <- g1

test_object <- readr::read_rds(sample)
# Make feature names, eg., "FT0004" for test
test <- test |> 
  rowid_to_column() |> 
  mutate(
    feature_id = paste0("FT", sprintf("%05d", row_number()))
  ) |> 
  select(feature_id, everything())


output_dir = "single_sample"
# Make the output dir if it does not exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

library(data.table)
test$rtime_adjusted <- test$rt
feature_peaks <- getBackendPeaks(test_object, fd = test)
spectra_df <- spectra(test_object)@backend@spectraData@listData |> bind_cols()
spectra_df$rtime_adjusted <- spectra_df$rtime
scan_info <- extractScanInfo(feature_peaks, fileNames(test_object), spectra_df)

specs <- 
  Ms2PseudoRaw_Final(
    scan_info,
    spectra(test_object),
    maxDiff = 0.01,
    output_dir = output_dir,
    return_spectra = TRUE
  )

# Open persistent DuckDB connection ONCE for the entire batch
con <- dbConnect(duckdb())
dbExecute(con, sprintf("CREATE OR REPLACE VIEW pointers AS SELECT * FROM read_parquet('%s/*_pointers.parquet')", output_dir))
dbExecute(con, sprintf("CREATE OR REPLACE VIEW pool AS SELECT * FROM read_parquet('%s/*_pool.parquet')", output_dir))
on.exit(dbDisconnect(con, shutdown = TRUE))

# First, find which features produced NAs:
feature_indices <- as.numeric(unique(scan_info$feature))
results <- batch_clean_spectra(feature_indices, output_dir, fetch_chunk_size = 100)
readr::write_rds(results, "single_sample/pseudo_ms2_unfiltered.rds")

results |> 
  filter(fdr<0.01) |> 
  write_rds("single_sample/pseudo_ms2_filtered.rds")

```






```{r load_dependencies}
#library(carpeDIAMS)
source("/faststorage/project/forensics_TOFscreenings/04_bbcid_ms2/carpeDIAMS/R/catch_ms2_functions.R")
source("/faststorage/project/forensics_TOFscreenings/04_bbcid_ms2/carpeDIAMS/R/cluster_functions.R")
source("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\carpeDIAMS\\R\\catch_ms2_functions.R")
source("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\carpeDIAMS\\R\\cluster_functions.R")


library(xcms)
library(dplyr)
library(tibble)
library(purrr)
library(readr)
library(ggplot2)
library(cowplot)
library(SummarizedExperiment)
#library(paletteer)
library(Spectra)
library(tidyr)

# Set global theme
theme_set(theme_minimal())
```

## Complete Tutorial

This tutorial demonstrates how to process DIA mass spectrometry data and create pseudo-DDA spectra using carpeDIAMS. We'll cover the complete workflow from data loading to visualization.

### 1. Data Loading and Preprocessing
First, we prepare our sample metadata and file paths. This step organizes the mzML files and creates sample groups for analysis.

```{r data_loading}
# Define file paths and create metadata
basepath <- "O:/HE_RKA-Forskning/data/metscreen_lind/data/02_pilot_rerun/mzML_files"
files <- list.files(basepath, recursive = TRUE, full.names = TRUE)
files <- files[grepl("[.]mzML", files)]
files <- files[!grepl("DDA|MS2|eV|K033|h5", files)]

# Create metadata tibble
metadata <- 
  tibble("filepath" = files,
         "group" = gsub(paste0(basepath, "/"), "", files)) |> 
  mutate(group = gsub("_.*", "", group))

# Filter to specific groups
metadata <- metadata |> filter(group %in% c("positive", "traffic"))
sample_files <- metadata$filepath

# Display metadata structure
print("Sample Metadata:")
head(metadata)
```

### 2. XCMS Processing Pipeline

This section demonstrates the standard XCMS workflow for processing DIA mass spectrometry data.

```{r xcms_processing}

if (file.exists("C:/Users/au557424/OneDrive - Aarhus universitet/Documents/object.rds")){
  object <- read_rds("C:/Users/au557424/OneDrive - Aarhus universitet/Documents/object.rds")
} else{
  # Create MsExperiment object
  object <- MsExperiment::readMsExperiment(
      spectraFiles = sample_files,
      backend = Spectra::MsBackendHdf5Peaks(),
      hdf5path = "hdf5",
      sampleData = metadata)
  
  # Filter retention time range (30-600 seconds)
  object <- ProtGenerics::filterSpectra(object, filterRt, rt = c(30, 600)) 
  
  # Peak detection using CentWave algorithm
  cwp <- CentWaveParam(ppm = 15,
                       peakwidth = c(2, 30),
                       prefilter = c(3, 500),
                       snthresh = 6,
                       mzdiff = 0.01,
                       mzCenterFun = "wMean",
                       integrate = 2)
  object <- findChromPeaks(object, param = cwp, chunkSize = 1, force.overwrite = TRUE)
  
  # Refine peaks by merging neighboring peaks
  mnpp <- MergeNeighboringPeaksParam(expandRt = 5)
  object <- refineChromPeaks(object, param = mnpp, chunkSize = 1)
  
  # First peak grouping (correspondence)
  pdp1 <- PeakDensityParam(sampleGroups = rep(1L, length(object)),
                          minFraction = 0.7,
                          binSize = 0.01,
                          ppm = 0,
                          bw = 30)
  object <- groupChromPeaks(object, param = pdp1)
  
  # Retention time adjustment
  pgp <- PeakGroupsParam(minFraction = 0.8, extraPeaks = 10, span = 0.5)
  object <- adjustRtime(object, param = pgp)
  
  # Second peak grouping after RT adjustment
  pdp2 <- PeakDensityParam(sampleGroups = rep(1L, length(object)),
                          minFraction = 0.5,
                          binSize = 0.01,
                          ppm = 0,
                          bw = 1.5)
  object <- groupChromPeaks(object, param = pdp2)
  
  # Fill missing peaks
  cpap <- ChromPeakAreaParam()
  object <- fillChromPeaks(object, param = cpap, chunkSize = 1)
  
  # Save and reload processed object
  write_rds(object, "object.rds")

}

print(paste("Processed", length(object), "samples"))
print(paste("Found", nrow(featureDefinitions(object)), "features"))
```


### 3. Data Structure Exploration

```{r}
object <- readr::read_rds("/faststorage/project/forensics_TOFscreenings/01_aging_study/new_xcms_workflow/output_annotation/tmp/grouped2.rds")
#object <- readr::read_rds("/faststorage/project/forensics_TOFscreenings/01_aging_study/new_xcms_workflow/output_all_samples/tmp/peak_integration3_dataset.rds")
```


#### Assay data

Let's examine the different data components: assay data, row data (features), and column data (samples).

```{r data_exploration}
library(SummarizedExperiment)
# Get quantification matrix (assay data)
quantified_data <- quantify(object)
assay_matrix <- assay(quantified_data)
head(assay_matrix)
write_csv(as_tibble(assay_matrix), "data/assay.csv")
```

#### Row data
```{r}
feature_info <- rowData(quantified_data)
head(feature_info)
write_csv(as_tibble(feature_info), "data/rowData.csv")
```

#### Sample data
```{r}
sample_data <- colData(quantified_data)
head(sample_data)
```





### 4. Feature Selection and Peak Analysis

Now we'll select features suitable for pseudo-DDA conversion and extract scan information.

```{r feature_selection}
fd <- featureDefinitions(object)
peaks <- xcms::chromPeaks(object)

mz_of_interest <- 271.2061
rt_of_interest <- 376.269

features_of_interest <- 
  fd |> tibble::as_tibble(rownames = "feature") |> 
  filter(abs(mzmed-mz_of_interest)<0.01,
         abs(rtmed-rt_of_interest)<20)

features_of_interest$mzmed
```

```{r}
param      <- xcms::PeakGroupsParam(
  smooth      = "loess",
  minFraction	= 0.8,
  span        = 1,
  extraPeaks  = 50,
  family      = "gaussian")

msobject <- adjustRtime(object, param)

pdp <- xcms::PeakDensityParam(sampleGroups = rep(1, length(msobject)), 
                              maxFeatures  = 50,
                              bw           = 2.5, 
                              minFraction  = 0.3,
                              binSize      = 0.01)
msobject <- groupChromPeaks(msobject, pdp)

feature_peaks <- getBackendPeaks(msobject, fd = features_of_interest)
filenames <- fileNames(msobject)
spectra_df <- spectra(msobject)@backend@spectraData@listData |> bind_cols()


```


```{r feature_selection}
# Get backend peaks for each feature
feature_peaks <- getBackendPeaks(object, fd = features_of_interest)
filenames <- fileNames(object)
spectra_df <- spectra(object)@backend@spectraData@listData |> bind_cols()
scan_info <- extractScanInfo(feature_peaks, filenames, spectra_df)
```

### 5. Pseudo-DDA Spectrum Generation

This section demonstrates how to convert DIA data to pseudo-DDA spectra by extracting MS2 information and applying correlation filtering.

```{r}
feature1 <- readr::read_rds("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\workflow\\parquets\\example_data\\feature1.rds")
source("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\carpeDIAMS\\R\\catch_ms2_functions.R")
source("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\carpeDIAMS\\R\\cluster_functions.R")
Rcpp::sourceCpp("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\workflow\\scripts\\gemini_upgma.cpp")

rowData <- readr::read_csv("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\workflow\\data\\rowData.csv")
fd <- readr::read_csv("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\workflow\\data\\feature_list.csv")
assay <- readr::read_csv("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\workflow\\data\\assay.csv")

# Phenylalanine: 796
# fd |> 
#   filter(abs(mzmed-166.0863)<0.001)

scan_info <- readr::read_csv("\\\\wsl.localhost\\Ubuntu-24.04\\home\\johan\\GenomeDK\\forensics_TOFscreenings\\04_bbcid_ms2\\workflow\\data\\scan_info.rds")
filenames <- colnames(assay)
feature1 |> head()
feature1$sample





```



##### Run a per sample clustering of mz values to generate sample-wise spectra that includes noise + random fragments
```{r pseudo_dda_generation}

# Show example of a single peak's scan information
single_sample <- scan_info |> group_split(sample) # We just select the first feature, if there are multiple, change the 1 accordingly to the index
#single_sample <- single_sample[1:10]
# Set parameters for pseudo-DDA conversion
minInt <- 1000  # Minimum intensity threshold
spectra_object <- spectra(msobject)
# Generate pseudo-raw MS2 spectra for each sample
ms2_specs <- 
  single_sample |> 
  map(~{
    Ms2PseudoRaw(
      .x,
      spectra_object,
      minInt = 200,
      maxDiff = 0.01
      )
    },
    .progress = TRUE) 

id = 1
filename <- fileNames(msobject)[id]

ms2 <- 
  seq_along(single_sample) |> 
  map(~{
    ss <- single_sample[[.x]] |> filter(feature == 796)
    Ms2PseudoRaw(
      filename = fileNames(msobject)[.x],
      acquisition_nums = ss |> pull(scan_indices),
      ms1Mz = ss$mz[1],
      apex_scan = ss$apex_scan_index[1],
      minInt = 200,
      maxDiff = 0.01,
      spectra_object=spectra_object
      )
  },.progress=T)




ms2 <- ms2 |> 
  map_dfr(~bind_rows(.x, .id = "feature"), .id = "sample")

ms2 <- ms2 |> mutate(feature = 796)
p <- 
  ms2 |> 
  filter(sample %in% subset_samples) |> 
  filter(mz > 120, mz < 120.1) |> 
  ggplot(aes(x=mz, y=log(int), color = factor(groups))) +
  geom_point()+
  facet_wrap(~sample, ncol = 1, scales = "free")+
  theme_minimal()

ggsave(plot = p,filename = "figures/p11.pdf", height = 20, width = 10, unit = "in", bg = "white")



ms2_specs[[1]] |> map_d

#ms2_specs_all <- ms2_specs

ms2_specs <- ms2_specs |> filter(feature == "1")
```

### 6. Visualization of Raw MS2 Data

Let's visualize the raw MS2 spectra before grouping to understand the data structure.

```{r raw_ms2_visualization, fig.height=6, fig.width=8}

ms2_specs$file <- sapply(ms2_specs$sample, function(x) filenames[as.numeric(x)])
# Visualize MS2 spectra from 4 random samples
sample_files <- sample(unique(ms2_specs$file), size = 4)
library(ggplot2)
ms2_raw_plot <- ms2_specs |>
  filter(file %in% sample_files) |> 
  ggplot(aes(x = mz, y = int, color = file)) +
  geom_line(show.legend = FALSE) +
  facet_wrap(~basename(file), scales = "free", ncol = 1) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Raw MS2 Spectra Across Samples",
       x = "m/z", 
       y = "Intensity") +
  theme_minimal() +
  theme(
    text = element_text(size = 8),
    plot.title = element_text(size = 8),
    axis.text = element_text(size = 7),
    strip.text = element_text(size = 7)
  )
ggsave(plot = ms2_raw_plot, filename = "figures/1.pdf", height =  5, width = 7, bg = "white")
print(ms2_raw_plot)
```

### 7. Peak Grouping and Correlation Analysis

This is the core step where we group similar m/z values across samples and calculate correlations.

```{r peak_grouping}
ms2_specs <- feature1
ms2_specs <- 
  ms2_specs |> 
  group_split(sample) |> 
  map(mutate, mz_g1 = fast_1D_upgma_grouping_cpp(fragment_mz, maxDiff=0.01),.progress=T)

ms2_specs <- 
  ms2_specs |>
  map_dfr(~{
    .x |> 
      mutate(file_scan = paste0(sample, scan)) |> 
  group_by(mz_g1) |> 
  mutate(mz_g1 = mean(fragment_mz)) |> 
  group_by(mz_g1, file_scan) |> 
  summarize(sample = sample[1],
            feature = feature[1],
            fragment_mz = mean(fragment_mz),
            int = sum(int),
            apex_scan = apex_scan[1],
            ms1_int = ms1_int[1],
            mz_g1 = mz_g1
            )
  },.progress=T)


# Add scan and file information for grouping
ms2_specs_extended <- 
  ms2 |> 
  group_by(groups, sample, scan) |> 
  mutate(mz_g1 = mean(mz)) |> 
  ungroup()


  
pseudo_groups <- fast_1D_upgma_grouping_cpp(ms2_specs_extended$mz, maxDiff=0.01)

# Group peaks with similar m/z values across all samples
pseudo_groups <- groupMultiple(
  mz = ms2_specs_extended$mz_g1, 
  int = ms2_specs_extended$int, 
  id = ms2_specs_extended$file_scan,
  maxDiff= 0.01
)


ms2_specs_extended <- ms2
ms2_specs_extended$grouped_mz <- pseudo_groups

ms2_specs_extended <- ms2_specs_extended |> group_by(grouped_mz) |> mutate(mz_g2 = mean(mz))

print(paste("Created", length(unique(pseudo_groups)), "m/z groups"))
```


```{r}
library(tidyverse)
library(xcms)
#ms2_specs_extended <-
# read_csv("../feature_M271.2061T376.269.csv")

#ms2_specs_extended <- ms2_specs_extended |> group_by(grouped_mz) |> mutate(mz_g2 = mean(mz))
ms2_specs_extended <- ms2_specs_extended |> filter(ms1_int > 1000)
ms2_specs_extended <- ms2_specs_extended |> ungroup() |> mutate(file = sapply(sample, FUN=function(x) filenames[x]))

consistent_groups <- 
  ms2_specs_extended |> 
  group_by(mz_g2) |> 
  summarize(count = n()) |> 
  filter(count > 500) |> 
  pull(mz_g2)

consistent_groups <- 
  ms2_specs_extended |> 
  filter(mz_g2 %in% consistent_groups) |> 
  group_by(sample, mz_g2) |>
  summarize(count = n()) |> 
  ungroup() |> 
  filter(count > 10) |> # Min elution count 
  group_by(mz_g2) |> 
  summarize(count = n()) |>  
  filter(count > 10) |> 
  pull(mz_g2)# Min number of samples with min elution counts

```


### 8. Correlation Analysis

Now we calculate elution and sample correlations to identify high-quality pseudo-DDA peaks.

```{r correlation_analysis}
# Calculate elution profile correlations (MS1 vs MS2 across retention time)
elution_correlation <- 
  ms2_specs_extended |> 
  filter(!is.na(ms1_int), mz_g2 %in% consistent_groups) |> 
  mutate(file_base = gsub("[.]mzML.*", "", basename(file)),
         g = paste0(file_base, grouped_mz)) |>
  group_by(g) |> 
  summarise(
    elution_cor = cor(log10(1 + int), log10(1 + ms1_int), use = "complete.obs"),
    ms1_total = sum(ms1_int, na.rm = TRUE),
    ms2_total = sum(int, na.rm = TRUE),
    n_scans = n(),
    mz_g2 = mz_g2[1],
    mz_g1 = mz_g1[1],
    mz = mean(mz, na.rm = TRUE),
    .groups = "drop")


library(data.table)

dt <- as.data.table(ms2_specs_extended)

# Pre-compute file_base én gang (undgå gsub per gruppe)
dt[, file_base := sub("[.]mzML.*", "", basename(file))]

# Brug setkey for hurtigere grouping
setkey(dt, file_base, mz_g2)

elution_correlation <- dt[
  !is.na(ms1_int) & mz_g2 %in% consistent_groups,
  {
    log_int <- log10(1 + int)
    log_ms1 <- log10(1 + ms1_int)
    .(
      elution_cor = cor(log_int, log_ms1, use = "complete.obs"),
      ms1_total = sum(ms1_int, na.rm = TRUE),
      ms2_total = sum(int, na.rm = TRUE),
      n_scans = .N,
      mz_grouped = mz_g2[1],
      mz_g1 = mz_g1[1],
      mz = mean(fragment_mz, na.rm = TRUE)
    )
  },
  by = .(file_base, mz_g2)
]


#
# elution_correlation |>  filter(n_scans > 5) |> arrange(-elution_cor) |> filter(grouped_mz == "G371") |> View()
# Calculate sample correlations (MS1 vs MS2 across samples)
sample_correlation <- 
  elution_correlation |>
  as_tibble() |> 
  filter(n_scans > 5) |> 
  select(grouped_mz = mz_grouped, everything()) |> 
  group_by(grouped_mz) |> 
  summarise(
    sample_cor = cor(log10(ms1_total + 1), log10(ms2_total + 1)),
    n_samples = n(),
    mz_g2 = mz_g2[1],
    mean_elution_cor = mean(elution_cor, na.rm = TRUE),
    mz = mz[1],
    .groups = "drop"
  ) |> 
  filter(n_samples > 5) 


# Combine correlation metrics
correlation_summary <- sample_correlation |> arrange(-mean_elution_cor)
print(paste("Correlation analysis completed for", nrow(correlation_summary), "groups"))
```

# Using mixed effects modeling with a negative binomial distribution

```{r}
library(glmmTMB)

sum(table(ms2_specs_extended$mz_g2)>100)

dat <- ms2_specs_extended |> filter(mz_g2 =="189.127338779994") 

extract_IRR <- function(coef_row, level = 0.95) {
  est <- coef_row["Estimate"]
  se  <- coef_row["Std. Error"]
  z   <- qnorm((1 + level) / 2)
  
  tibble(
    IRR      = exp(est),
    CI_lower = exp(est - z * se),
    CI_upper = exp(est + z * se)
  )
}



# Example: one predictor x, with random intercept for Block
start <- Sys.time()
IRR <- ms2_specs_extended |>
  group_by(mz_g2) |> 
  mutate(count = n()) |> 
  ungroup() |> 
  filter(count>100) |> 
  group_split(mz_g2, .keep = T) |> 
  map_dfr(~{
    glmmTMB(
      ms1_int ~ int + (1 | file),
      family = nbinom2(),         # use nbinom1() if appropriate
      data = .x) |>
      summary() |> 
      coef() |> 
      (\(x) x$cond["int", ])() |>
      extract_IRR() |> 
      mutate(mz_g2 = .x$mz_g2[1])},
      .progress=T
    )

GausEstimate <- ms2_specs_extended |>
  group_by(mz_g2) |> 
  mutate(count = n(), int = scale(int), ms1_int = scale(ms1_int)) |> 
  ungroup() |> 
  filter(count>1000) |> 
  group_split(mz_g2, .keep = T) |> 
  map_dfr(~{
    glmmTMB(
      ms1_int ~ int + (1 | file),
      family = gaussian(),         # use nbinom1() if appropriate
      data = .x) |>
      summary() |> 
      coef() |> 
      (\(x) x$cond["int", "Estimate"])() |> 
      as_tibble() |> 
      mutate(mz_g2 = .x$mz_g2[1])},
      .progress=T
    )

mz_g2_overall_cor <- 
   ms2_specs_extended |>
  group_by(mz_g2) |> 
  summarize(
    correlation = cor(int, ms1_int, method = "spearman")
  ) |> 
  arrange(-correlation)

mz_g2_overall_cor.p <- 
   ms2_specs_extended |>
  group_by(mz_g2) |> 
  summarize(
    p = cor(int, ms1_int, method = "pearson")
  ) |> 
  arrange(-p)
  
mz_g2_IRR <- 
  IRR |> 
  bind_rows() |> 
  arrange(-IRR)
  
mz_g2_Coef <- 
  GausEstimate |> 
  arrange(-value)

```


```{r}
correlation_summary <- 
  sample_correlation |> 
  left_join(mz_g2_IRR) |> 
  left_join(mz_g2_Coef) |> 
  left_join(mz_g2_overall_cor) |> 
  left_join(mz_g2_overall_cor.p)

correlation_summary |> 
  ggplot(aes(x=sample_cor, y=mean_elution_cor, color = p, size = p))+
  geom_point()+
  scale_color_distiller(palette = "Spectral") +
  labs(
    title = "MS2 Peak Quality Assessment",
    x = "Sample Correlation (MS1 vs MS2)",
    y = "Mean Elution Correlation (MS1 vs MS2)",
    caption = "Points in upper right represent high-quality peaks",
    color = "Total Pearson r",
    size = "Total Pearson r"
  ) +
  theme_classic() +
  theme(
    text = element_text(size = 8),
    plot.title = element_text(size = 8),
    axis.text = element_text(size = 7),
    plot.caption = element_text(size = 6)
  )

```


### 9. Quality Control Visualization

Let's visualize the correlation patterns to assess data quality.

```{r correlation_visualization, fig.height=4, fig.width=6}
correlation_plot <- correlation_summary |> 
  ggplot(aes(x = sample_cor, y = mean_elution_cor)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  labs(
    title = "MS2 Peak Quality Assessment",
    x = "Sample Correlation (MS1 vs MS2)",
    y = "Mean Elution Correlation (MS1 vs MS2)",
    caption = "Points in upper right represent high-quality peaks"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 8),
    plot.title = element_text(size = 8),
    axis.text = element_text(size = 7),
    plot.caption = element_text(size = 6)
  )

print(correlation_plot)
```

### 10. High-Quality Peak Selection and Detailed Analysis

Now we'll select high-quality peaks based on correlation thresholds and perform detailed analysis.

```{r peak_selection}
# Define quality thresholds
cor_threshold <- 0.5

# Select high-quality peaks
high_quality_peaks <- correlation_summary |> 
  filter(p>cor_threshold) |>
  filter(n_samples>5)

print(paste("Selected", nrow(high_quality_peaks), "high-quality peaks"))

# Select a specific peak for detailed visualization
if(nrow(high_quality_peaks) > 0) {
  selected_group <- high_quality_peaks$grouped_mz[1]  # Select 4th best peak
  selected_mz <- high_quality_peaks$mz[1]
  
  print(paste("Analyzing peak at m/z =", round(selected_mz, 4)))
}
```

### 11. Detailed Peak Visualization

Let's create comprehensive visualizations for the selected high-quality peak.

** note that you can chance selected_group to whatever group of mz values you would like to visualize **

```{r detailed_visualization, fig.height=8, fig.width=10}
library(zoo)

if(exists("selected_group")) {
  # Extract data for selected peak
  selected_peak_data <- 
    ms2_specs_extended |> 
    filter(grouped_mz == selected_group) |> 
    mutate(
      file_base = gsub("[.]mzML.*", "", basename(file)),
      ms1_int = replace_na(ms1_int, 0)
    ) |> 
    group_by(file_base) |>
    mutate(
      log_ms1 = log10(mean(ms1_int) + 1),
      scan = scan-apex_scan,
      ms1_int = zoo::rollmean(ms1_int, k = 2, fill = NA),
      int = zoo::rollmean(int, k = 2, fill = NA)
      )
  
  # Elution profile plot (MS2)
  elution_ms2 <- selected_peak_data |>
    ggplot(aes(x = scan, y = int, color = log_ms1, group = file_base)) +
    #geom_line(show.legend = FALSE) +
    geom_point()+
    scale_color_gradientn(colors = paletteer::paletteer_c("ggthemes::Blue-Green Sequential", 30)) +
    #scale_y_continuous(trans = "log10") +
    labs(title = "MS2 Elution Profile", y = "MS2 Intensity") +
    theme_minimal() +
    theme(
      text = element_text(size = 8),
      plot.title = element_text(size = 8),
      axis.text = element_text(size = 7),
      axis.text.x = element_blank(),
      axis.title.x = element_blank()
    )
  
  # Elution profile plot (MS1)
  elution_ms1 <- selected_peak_data |>
    ggplot(aes(x = scan, y = ms1_int, group = file_base)) +
    geom_line(color = "gray60") +
    labs(title = "MS1 Elution Profile", y = "MS1 Intensity", x = "Scan Number") +
    theme_minimal() +
    theme(
      text = element_text(size = 8),
      plot.title = element_text(size = 8),
      axis.text = element_text(size = 7)
    )
  
  # Sample correlation plot
  sample_corr_data <- selected_peak_data |> 
    group_by(file_base) |> 
    summarise(
      ms2_intensity = sum(int),
      ms1_intensity = sum(ms1_int),
      .groups = "drop"
    )
  
  sample_corr_plot <- sample_corr_data |>
    ggplot(aes(x = ms1_intensity, y = ms2_intensity)) +
    geom_point(color = "steelblue", alpha = 0.7) +
    scale_x_continuous(trans = "log10") + 
    scale_y_continuous(trans = "log10") +
    labs(
      title = "Sample Correlation",
      x = "MS1 Intensity", 
      y = "MS2 Intensity"
    ) +
    theme_minimal() +
    theme(
      text = element_text(size = 8),
      plot.title = element_text(size = 8),
      axis.text = element_text(size = 7)
    )
  
  # Combine elution plots
  combined_elution <- plot_grid(elution_ms2, elution_ms1, 
                               ncol = 1, rel_heights = c(3, 2), 
                               align = "hv")
  
  # Final combined plot
  final_plot <- plot_grid(combined_elution, sample_corr_plot, 
                         ncol = 2, rel_widths = c(2, 1))
}
print(final_plot)
```

### 12. Pseudo-DDA Spectrum Creation

Finally, let's create the consolidated pseudo-DDA spectrum from high-quality peaks.

```{r pseudo_spectrum, fig.height=5, fig.width=8}
# Create consolidated spectrum from high-quality peaks
if(exists("high_quality_peaks") && nrow(high_quality_peaks) > 0) {
  
  pseudo_spectrum <- ms2_specs_extended |> 
    filter(grouped_mz %in% high_quality_peaks$grouped_mz) |> 
    group_by(grouped_mz) |> 
    summarise(
      mz = mean(mz),
      total_intensity = sum(int),
      .groups = "drop"
    ) |> 
    arrange(mz)
  
  # Create spectrum plot
  spectrum_plot <- pseudo_spectrum |>
    ggplot(aes(x = mz, y = total_intensity)) +
    geom_segment(aes(x = mz, xend = mz, y = 0, yend = total_intensity), 
                 color = "steelblue", size = 0.8) +
    labs(
      title = "Consolidated Pseudo-DDA Spectrum",
      x = "m/z",
      y = "Summed Intensity",
      caption = paste("Spectrum contains", nrow(pseudo_spectrum), "high-quality peaks")
    ) +
    theme_minimal() +
    theme(
      text = element_text(size = 8),
      plot.title = element_text(size = 8),
      axis.text = element_text(size = 7),
      plot.caption = element_text(size = 6)
    )
  
  print(spectrum_plot)
  
  # Summary statistics
  cat("\n=== PSEUDO-DDA SUMMARY ===\n")
  cat("Total peaks in pseudo-spectrum:", nrow(pseudo_spectrum), "\n")
  cat("m/z range:", round(min(pseudo_spectrum$mz), 2), "-", round(max(pseudo_spectrum$mz), 2), "\n")
  cat("Intensity range:", round(min(pseudo_spectrum$total_intensity), 0), "-", 
      round(max(pseudo_spectrum$total_intensity), 0), "\n")
}

```


#### save the pseudo spectrum to mgf (not tested!)

```{r}
library(Spectra)

# Create a Spectra object
spd <- DataFrame(
  msLevel = 2L,
  rtime = 376.269, #features_of_interest$rtmed,
  precursorMz = 271.2061# features_of_interest$mzmed  # Add if you know it
)

spd$mz <- list(pseudo_spectrum$mz)
i <- pseudo_spectrum$total_intensity
spd$intensity <- list(1000*(i-min(i))/(max(i)-min(i)))

spec <- Spectra(spd)

# Export to MGF
export(spec, backend = MsBackendMzR(), file = "C:/Users/au557424/Aarhus universitet/output.mgf")
```



## Protocol Summary

The **carpeDIAMS** workflow follows these key steps:

1. **Data Loading**: Import mzML files and create sample metadata
2. **XCMS Processing**: Standard peak detection, grouping, and retention time alignment
3. **Feature Selection**: Identify suitable features for pseudo-DDA conversion
4. **MS2 Extraction**: Extract MS2 information for selected features
5. **Peak Grouping**: Group similar m/z values across samples using agglomerative clustering
6. **Correlation Analysis**: Calculate elution and sample correlations to assess peak quality
7. **Quality Filtering**: Select high-quality peaks based on correlation thresholds
8. **Spectrum Consolidation**: Create final pseudo-DDA spectra from filtered peaks

## Key Parameters

- **minInt**: Minimum intensity threshold for MS2 peaks (default: 500)
- **Elution correlation threshold**: Minimum correlation between MS1 and MS2 elution profiles (default: 0.3)
- **Sample correlation threshold**: Minimum correlation between MS1 and MS2 across samples (default: 0.8)
- **m/z tolerance**: Tolerance for grouping similar m/z values (default: 0.001 Da)

This approach enables the extraction of pseudo-DDA spectra from DIA data while maintaining high confidence through correlation-based quality control.


```{r}

sps <- Spectra(backend = MsBackendMsp::MsBackendMsp(), 
               file = "MS_library - LC-MS pos.msp")

sps <- sps[unique(sps$Name) |> sapply(FUN = \(x){which(x==sps$Name)[1]})]
```



# Full pipeline
```{r}

fd <- featureDefinitions(object)

start <- Sys.time()

spectra <-
  seq_along(sps$Name) |> 
  map(~{
  
  example <- sps[sps$Name == sps$Name[.x]]
  
  mz_of_interest <- example@backend@spectraData$precursorMz
  rt_of_interest <- as.numeric(example@backend@spectraData$RT)*60
  
  features_of_interest <- 
    fd |> as_tibble(rownames = "feature") |> 
    filter(abs(mzmed-mz_of_interest[1])<0.001,
           abs(rtmed-rt_of_interest[1])<10)
  
  if (nrow(features_of_interest)==0){
    return("No peak found in MS1")
  }
  
  # Get backend peaks for each feature
  feature_peaks <- getBackendPeaks(object, fd = features_of_interest)
  filenames <- fileNames(object)
  spectra_df <- spectra(object)@backend@spectraData@listData |> bind_cols()
  
  # Extract scan information for selected features
  scan_info <- feature_peaks |> map(extractScanInfo, filenames, spectra_df, .progress = TRUE)
  
  # Show example of a single peak's scan information
  single_peak <- scan_info[[1]] # We just select the first feature, if there are multiple, change the 1 accordingly to the index
  
  # Set parameters for pseudo-DDA conversion
  minInt <- 1000  # Minimum intensity threshold
  
  # Generate pseudo-raw MS2 spectra for each sample
  ms2_specs <- 
    single_peak |> 
    map(Ms2PseudoRaw, minInt = minInt, maxDiff = 0.001) |> 
    set_names(single_peak |> map_chr(~.x$filename)) |> 
    bind_rows(.id = "file")
  
  # Add scan and file information for grouping
  ms2_specs_extended <- 
    ms2_specs |> 
    mutate(file_scan = paste0(basename(file))) |> 
    group_by(groups, file_scan) |> 
    mutate(mz_g1 = mean(mz)) |> 
    ungroup()
  
  
  # Group peaks with similar m/z values across all samples
  ms2_specs_extended$grouped_mz <- groupMultiple(
    mz = ms2_specs_extended$mz_g1, 
    int = ms2_specs_extended$int, 
    id = ms2_specs_extended$file_scan,
    maxDiff = 0.01
  )
  
  ms2_specs_extended <- ms2_specs_extended |> group_by(grouped_mz) |> mutate(mz_g2 = mean(mz))
  
  # Calculate elution profile correlations (MS1 vs MS2 across retention time)
  elution_correlation <- 
    ms2_specs_extended |> 
    filter(!is.na(ms1_int)) |> 
    mutate(file_base = gsub("[.]mzML.*", "", basename(file))) |>
    group_by(grouped_mz, file_base) |> 
    summarise(
      elution_cor = cor(log10(1 + int), log10(1 + ms1_int), use = "complete.obs"),
      ms1_total = sum(ms1_int, na.rm = TRUE),
      ms2_total = sum(int, na.rm = TRUE),
      n_scans = n(),
      mz_g2 = mz_g2[1],
      mz_g1 = mz_g1[1],
      mz = mean(mz, na.rm = TRUE),
      .groups = "drop"
    )
  
  #
  
  # Calculate sample correlations (MS1 vs MS2 across samples)
  correlation_summary <- 
    elution_correlation |>
    filter(n_scans > 5) |> 
    group_by(grouped_mz) |> 
    summarise(
      sample_cor = cor(log10(ms1_total + 1), log10(ms2_total + 1)),
      n_samples = n(),
      mean_elution_cor = mean(elution_cor, na.rm = TRUE),
      mz = mz[1],
      .groups = "drop"
    )
  
  elution_threshold <- 0.8
  sample_threshold <- 0.8
  
  # Select high-quality peaks
  high_quality_peaks <- correlation_summary |> 
    filter(mean_elution_cor > elution_threshold | sample_cor > sample_threshold) |> 
    arrange(desc(mean_elution_cor))
  
  print(paste("Selected", nrow(high_quality_peaks), "high-quality peaks"))
  
  # Select a specific peak for detailed visualization
  if(nrow(high_quality_peaks) > 0) {
    selected_group <- high_quality_peaks$grouped_mz[3]  # Select 4th best peak
    selected_mz <- high_quality_peaks$mz[3]
    
    pseudo_spectrum <- ms2_specs_extended |> 
      filter(grouped_mz %in% high_quality_peaks$grouped_mz) |> 
      group_by(grouped_mz) |> 
      summarise(
        mz = mean(mz),
        total_intensity = sum(int),
        .groups = "drop"
      ) |> 
      arrange(mz)
    print(paste("Analyzing peak at m/z =", round(selected_mz, 4)))
  }
  
  
  # Create a Spectra object
  spd <- DataFrame(
    msLevel = 2L,
    rtime = features_of_interest$rtmed,
    compound_name = sps$Name[.x],
    precursorMz = features_of_interest$mzmed  # Add if you know it
  )
  
  spd$mz <- list(pseudo_spectrum$mz)
  i <- pseudo_spectrum$total_intensity
  spd$intensity <- list(1000*(i-min(i))/(max(i)-min(i)))
  
  return(Spectra(spd))
}, .progress=T)
end <- Sys.time()

print(paste0("time used: ", end-start))


export(spec, backend = MsBackendMgf::MsBackendMgf(), file = "my_spectra.mgf")
spectra |> map_lgl(is.character) |> mean()

scores <- c()
counts <- c()
plots <- list()
for (id in seq_along(spectra)){
  if (is.character(spectra[[id]])){
    scores <- c(scores, NA)
    counts <- c(counts, NA)
    plots[[length(plots)+1]] <- NA
    next
    }
  #scores <- c(scores, compareSpectra(spectra[[id]][1], sps[id], fun = MsCoreUtils::ndotProduct()))
  scores <- c(scores, gnpsScore(query = spectra[[id]][1], target = sps[id], tolerance = 0.01)$score)
  counts <- c(counts, gnpsScore(query = spectra[[id]][1], target = sps[id], tolerance = 0.01)$n_matches)
  }

tibble(scores, counts) |> ggplot(aes(x=scores, y=counts))+geom_point()
tibble(scores, counts) |> filter(!is.na(scores)) |> filter(scores > 0.6, counts > 4) |> ggplot(aes(x=scores, y=counts))+geom_point()
plotSpectraMirror(spectra[[id]], sps[id],
     ppm = 50, main = "dia vs dda", print=F)
score <- compareSpectra(spectra[[id]], sps[id], fun = MsCoreUtils::dotProduct)


for (i in seq_along(spectra)) {
  if (is.character(spectra[[i]])) next
  spectra[[i]]$adduct <- "[M+H]+"
  spectra[[i]]$ionMode <- "positive"
  spectra[[i]]$charge <- 1L
  spectra[[i]]$title <- paste0("Compound_", i)
}

combined_spectra <- do.call(c, spectra[!(spectra |> map_lgl(is.character))])
export(combined_spectra, backend = MsBackendMgf::MsBackendMgf(), file = "my_spectra.mgf")

sps$Name[31]

plotSpectraMirror(spectra[[31]][1], sps[31],
     ppm = 50, main = "dia vs dda")

```


```{r}
library(Spectra)

gnpsScore <- function(query, target, tolerance = 0.02) {
  mz_q <- mz(query) |> unlist()
  int_q <- intensity(query)  |> unlist()
  
  mz_t <- mz(target) |> unlist()
  int_t <- intensity(target) |> unlist()
  
  # Find matches: for hver peak i query, find nærmeste peak i target
  matched_q <- c()
  matched_t <- c()
  
  for (i in seq_along(mz_q)) {
    diffs <- abs(mz_t - mz_q[i])
    min_diff <- min(diffs)
    if (min_diff <= tolerance) {
      j <- which.min(diffs)
      matched_q <- c(matched_q, i)
      matched_t <- c(matched_t, j)
    }
  }
  
  if (length(matched_q) == 0) {
    return(data.frame(
      n_matches = 0,
      score = NA_real_
    ))
  }
  
  int_q_matched <- int_q[matched_q]
  int_t_matched <- int_t[matched_t]
  
  # Cosine similarity
  score <- sum(int_q_matched * int_t_matched) /
           (sqrt(sum(int_q_matched^2)) * sqrt(sum(int_t_matched^2)))
  
  return(data.frame(
    n_matches = length(matched_q),
    score = score
  ))
}




n = 1
gnpsScore(query = spectra[[great_matches[n]]][1], target = sps[great_matches[n]], tolerance = 0.01)
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[great_matches[n]],
     ppm = 50, main = "dia vs dda")
sps$Name[great_matches[n]]
n = n+1
```


```{r}
sps <- Spectra(backend = MsBackendMsp::MsBackendMsp(), 
               file = "MoNA-export-LC-MS-MS_Positive_Mode.msp")

n=8
great_matches <- which(((counts>3)&(counts<10)&(scores>0.6)))
match_info <- 
  seq_along(sps) |> 
  map_dfr(
   ~{gnpsScore(query = spectra[[great_matches[n]]][1], target = sps[.x], tolerance = 0.01)} 
  )
  
top_matches <- 
  match_info |>
  rowid_to_column() |> 
  mutate(precurserMz = sps[great_matches[n]]@backend@spectraData@listData$precursorMz,
         matchPrecurserMz = sps[rowid]@backend@spectraData@listData$precursorMz) |> 
  filter(n_matches > 2, score > 0.6, abs(precurserMz-matchPrecurserMz)<0.01) |> 
  arrange(-n_matches, -score) |> 
  slice_head(n=10) 

great_matches[n]
sps$Name[great_matches[n]]

sps$Name[top_matches$rowid[1]]
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[top_matches$rowid[1]],
     ppm = 50, main = "dia vs dda")
sps$Name[top_matches$rowid[2]]
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[top_matches$rowid[2]],
     ppm = 50, main = "dia vs dda")
sps$Name[top_matches$rowid[3]]
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[top_matches$rowid[3]],
     ppm = 50, main = "dia vs dda")
sps$Name[top_matches$rowid[4]]
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[top_matches$rowid[4]],
     ppm = 50, main = "dia vs dda")


sps[great_matches[n]]@backend@spectraData
sps[top_matches$rowid[1]]@backend@spectraData
sps[536]@backend@spectraData

spectra[[26]]@backend@spectraData

```




```{r}
sps <- Spectra(backend = MsBackendMsp::MsBackendMsp(), 
               file = "MoNA-export-LC-MS-MS_Positive_Mode.msp")

fl <- "MoNA-export-LC-MS-MS_Positive_Mode.sdf"
cmps <- compound_tbl_sdf(fl)
metad <- CompoundDb::make_metadata(source = "MoNA", source_version = "0.3",
    source_date = NA, organism = "Hsapiens", url = "https://mona.fiehnlab.ucdavis.edu/downloads")
db_f <- createCompDb(cmps, metadata = metad, path = tempdir())

db_file <- createCompDb(cmps, metadata = metad, overwrite=T,
                        path = tempdir())


n=8
great_matches <- which(((counts>10)&(scores>0.6)))
match_info <- 
  seq_along(sps) |> 
  map_dfr(
   ~{gnpsScore(query = spectra[[23]][1], target = sps[.x], tolerance = 0.01)},.progress=T
  )
  
top_matches <- 
  match_info |>
  rowid_to_column() |> 
  mutate(precurserMz = sps[great_matches[n]]@backend@spectraData@listData$precursorMz,
         matchPrecurserMz = sps[rowid]@backend@spectraData@listData$precursorMz) |> 
  filter(n_matches > 2, score > 0.6, abs(precurserMz-matchPrecurserMz)<0.01) |> 
  arrange(-n_matches, -score) |> 
  slice_head(n=10) 

great_matches[n]
sps$Name[great_matches[n]]

sps$Name[top_matches$rowid[1]]
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[top_matches$rowid[1]],
     ppm = 50, main = "dia vs dda")
sps$Name[top_matches$rowid[2]]
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[top_matches$rowid[2]],
     ppm = 50, main = "dia vs dda")
sps$Name[top_matches$rowid[3]]
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[top_matches$rowid[3]],
     ppm = 50, main = "dia vs dda")
sps$Name[top_matches$rowid[4]]
plotSpectraMirror(spectra[[great_matches[n]]][1], sps[top_matches$rowid[4]],
     ppm = 50, main = "dia vs dda")


sps[great_matches[n]]@backend@spectraData
sps[top_matches$rowid[1]]@backend@spectraData
sps[536]@backend@spectraData

spectra[[26]]@backend@spectraData

```



```{r}
library(httr)
library(jsonlite)

usi <- "mzspec:GNPS:GNPS-LIBRARY:accession:CCMSLIB00005436077"
url <- paste0("https://metabolomics-usi.gnps2.org/api/spectrum?usi=", URLencode(usi))
url <- "https://metabolomics-usi.gnps2.org/json/?usi1=mzspec%3AGNPS%3AGNPS-LIBRARY%3Aaccession%3ACCMSLIB00005436077"
url <- "mzspec:MSV000089306:raw/Phoenicin switch 2/JVC00907.mzML:scan:4294967295"

base_api <- "https://metabolomics-usi.gnps2.org/json/?usi1="


https://metabolomics-usi.gnps2.org/json/?usi1=mzspec%3AGNPS%3AGNPS-LIBRARY%3Aaccession%3ACCMSLIB00005436077
# Your USI
usi <- "mzspec:MSV000089306:raw/Phoenicin switch 2/JVC00907.mzML:scan:4294967295"

# Build the API URL
url <- paste0("https://metabolomics-usi.gnps2.org/json/?usi=", URLencode(usi))

# Fetch the spectrum
res <- GET(url)

# Parse the JSON response
spec_data <- fromJSON(content(res, "text"))

# Access m/z and intensity
mz <- spec_data$peaks[,1]
intensity <- spec_data$peaks[,2]

# Create a Spectra object
spd <- DataFrame(
  msLevel = 2L,
  rtime = NA_real_,
  compound_name = NA_character_,
  precursorMz = spec_data$precursor_mz  # Add if you know it
)

spd$mz <- list(mz)
i <- intensity
spd$intensity <- list(1000*(i-min(i))/(max(i)-min(i)))


# Create a Spectra object
spd <- DataFrame(
  msLevel = 2L,
  rtime = features_of_interest$rtmed,
  precursorMz = features_of_interest$mzmed  # Add if you know it
)

spd$mz <- list(pseudo_spectrum$mz)
i <- pseudo_spectrum$total_intensity
spd$intensity <- list(1000*(i-min(i))/(max(i)-min(i)))

spec <- Spectra(spd)
```


